---
title: ''
output: html_document
date: "2023-10-10"
---

v1
Intended to data shuffling, splitting for bootstrapping for los
XGB fails at loading. Dmatrix requires to be done outside of the list


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning = F,results = T)
rm(list = ls())
library(tidyverse)
library(patchwork)
library(rstanarm)
library(bayesplot)
library(loo)
library(lme4)
library(gtsummary)
library(fairness)
library(caret)
library(randomForest)
library(xgboost)

cbPalette <- c("#E64B35", "#4DBBD5")
source("../allofus_los_Jtrans2024/helpers.R")
load("../allofus_los_Jtrans2024/readmissionProject/asthma.read.multivariate.full.stan.v7.RData")

rm(glmer.read.ri.full.d,glmer.read.ri.full.e,glmer.read.ri.full.f,glmer.read.ri.full.g,
   stan.read.ri.full.d,stan.read.ri.full.e,stan.read.ri.full.f,stan.read.ri.full.g,t_prior,coefs,knots,cbPalette)
```





With patient-level partitioning, there is not cases classified as readmitted ('Yes') due to probability lower than 0.5. We should test this partitioning at visits-level.


# Dataset partitioning by visits-level

```{r}
data.splitted<-list()
```


```{r}
myLevels<-c('Yes','No')
# visit occurrence is not unique
# data$visit_id<-as.character(data$visit_occurrence_id)
# data %>% 
#   summarise(count=n_distinct(visit_id))

#let's create it
data2<-data %>% 
  group_by(id) %>% 
  arrange(visit_start_date) %>% 
  mutate(id_visit=paste0(id,'-',row_number()),
         losBinaryNum=ifelse(los==0,0,1),
         losBinaryFactor=factor(ifelse(losBinaryNum==0,"ShortStay","ProlongedStat"),levels=c("ShortStay","ProlongedStat")))%>% 
  ungroup()
#check proportions on data2
data2 %>% 
  group_by(race3,readmission30d) %>% 
  summarise(n=n()) %>% 
  mutate(prop=(100*n/nrow(data2)) %>% round(1),
         prop2t=paste0(paste0(n," (",prop,")"))
         ) %>% 
  select(race3,readmission30d,prop2t)
data2$readmission30d<-factor(data2$readmission30d,levels=myLevels)

#check proportions on data2
data2 %>% 
  group_by(race3,losBinaryFactor) %>% 
  summarise(n=n()) %>% 
  mutate(prop=(100*n/nrow(data2)) %>% round(1),
         prop2t=paste0(paste0(n," (",prop,")"))
         ) %>% 
  select(race3,losBinaryFactor,prop2t)
data2$readmission30d<-factor(data2$readmission30d,levels=myLevels)

summary(data2$los)
```



```{r}
# print("cp upsampling")
m.age <- glmer(
  losBinaryNum ~
    s.age+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  nAGQ = 0,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.age<-tbl_regression(m.age,exp=T)
```


```{r}
m.sex <- glmer(
  losBinaryNum ~
    sex_at_birth+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  nAGQ = 0,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.sex<-tbl_regression(m.sex,exp=T)
```


```{r}
m.dep <- glmer(
  losBinaryNum ~
    s.dep+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  nAGQ = 0,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.dep<-tbl_regression(m.dep,exp=T)
t.dep

```


```{r}
m.hct <- glmer(
  losBinaryNum ~
    hcType1+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  nAGQ = 0,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.hct<-tbl_regression(m.hct,exp=T)
t.hct
```


```{r}
m.race <- glmer(
  losBinaryNum ~
    race3+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  nAGQ = 0,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.race<-tbl_regression(m.race,exp=T)
```


```{r}
m.week <- glmer(
  losBinaryNum ~
    weekend+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  # nAGQ = 0,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.week<-tbl_regression(m.week,exp=T)
```


```{r}
m.seas <- glmer(
  losBinaryNum ~
    season+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  # nAGQ = 0,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)
t.seas<-tbl_regression(m.seas,exp=T)
tbl_regression(m.seas,exp=T)

m.seas2 <- glmer(
  readmission30dNum ~
    season+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  # nAGQ = 0,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)
tbl_regression(m.seas2,exp=T)
```

```{r}
m.AD <- glmer(
  losBinaryNum ~
    AD.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  # nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.ad<-tbl_regression(m.AD,exp=T)
```


```{r}
m.AR <- glmer(
  losBinaryNum ~
    AR.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  # nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.ar<-tbl_regression(m.AR,exp=T)
```



```{r}
m.cancer <- glmer(
  losBinaryNum ~
    Cancer.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  # nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.cancer<-tbl_regression(m.cancer,exp=T)
```


```{r}
m.chd <- glmer(
  losBinaryNum ~
    CHD.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.chd<-tbl_regression(m.chd,exp=T)
```


```{r}
m.ckd <- glmer(
  losBinaryNum ~
    CKD.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.ckd<-tbl_regression(m.ckd,exp=T)
```


```{r}
m.copd <- glmer(
  losBinaryNum ~
    COPD.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.copd<-tbl_regression(m.copd,exp=T)
```


```{r}
m.depr<- glmer(
  losBinaryNum ~
    Depression.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.depr<-tbl_regression(m.depr,exp=T)
```



```{r}
m.diab <- glmer(
  losBinaryNum ~
    Diab.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  # nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.diab<-tbl_regression(m.diab,exp=T)
```


```{r}
m.eoe <- glmer(
  losBinaryNum ~
    EoE.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.eoe<-tbl_regression(m.eoe,exp=T)
```


```{r}
m.fa <- glmer(
  losBinaryNum ~
    FA.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  # nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.fa<-tbl_regression(m.fa,exp=T)
```


FA not significant

```{r}
m.gerd <- glmer(
  losBinaryNum ~
    GERD.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.gerd<-tbl_regression(m.gerd,exp=T)
```



```{r}
m.hta <- glmer(
  losBinaryNum ~
    HTA.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  # nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.hta<-tbl_regression(m.hta,exp=T)
```


```{r}
m.obst <- glmer(
  losBinaryNum ~
    OBST.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  # nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.obst<-tbl_regression(m.obst,exp=T)
```


```{r}
m.apnea <- glmer(
  losBinaryNum ~
    APNEA.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  # nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.apnea<-tbl_regression(m.apnea,exp=T)
```




```{r}
m.psor <- glmer(
  losBinaryNum ~
    PSOR.lon+
    (1 | id) + (1 | ehrsite),
  data = data2,
  family = binomial(),
  # nAGQ = 1,
  glmerControl(
    optimizer = "bobyqa",
    calc.derivs = FALSE,
    optCtrl = list(maxfun = 2e5)
  )
)

t.psor<-tbl_regression(m.psor,exp=T)
```


Psoriasis is not significant

```{r}
tbl_stack(list(t.age,t.sex, t.race,t.dep,t.week,t.seas))
# FA, EoE and PSOR are no significant
tbl_stack(list(t.ad,t.ar, t.cancer,t.chd,t.ckd,t.copd,t.depr,t.diab,t.eoe,t.fa,t.gerd,t.hta,t.obst,t.apnea,t.psor))
```


no eoe, no fa, no psor
