---
title: ''
output: html_document
date: "2023-10-10"
---

v1
bootstrap for random forest (race-based and race-neutral), includes no sampling an oversampling 
code clean-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning = F,results = T)
rm(list = ls())
library(tidyverse)
library(patchwork)
library(rstanarm)
library(bayesplot)
library(loo)
library(lme4)
library(gtsummary)
library(fairness)
library(caret)
library(randomForest)
library(xgboost)

cbPalette <- c("#E64B35", "#4DBBD5")
source("../allofus_los_Jtrans2024/helpers.R")
source("myFairness.R")

myLevels<-c("ProlongedStay","ShortStay")
load("data.splitted.los.v1.RData")
```


```{r}
rb.eq<-formula(losBinaryFactor~
                 s.age +
                 sex_at_birth +
                 s.dep +
                 race3 +
                 hcType1 +
                 weekend +
                 season +
                 AD.lon +
                 AR.lon +
                 Cancer.lon +
                 CHD.lon +
                 CKD.lon +
                 COPD.lon +
                 Depression.lon +
                 Diab.lon +
                 # FA.lon +
                 # EoE.lon
                 GERD.lon +
                 HTA.lon +
                 OBST.lon +
                 # PSOR.lon +
                 APNEA.lon)

rn.eq<-formula(losBinaryFactor~
                 s.age +
                 sex_at_birth +
                 s.dep +
                 # race3 +
                 hcType1 +
                 weekend +
                 season +
                 AD.lon +
                 AR.lon +
                 Cancer.lon +
                 CHD.lon +
                 CKD.lon +
                 COPD.lon +
                 Depression.lon +
                 Diab.lon +
                 # FA.lon +
                 # EoE.lon
                 GERD.lon +
                 HTA.lon +
                 OBST.lon +
                 # PSOR.lon +
                 APNEA.lon)
```


### Random forest no upsampling

```{r}
model1Label="rb.rf"
model2Label="rn.rf"
upsampling=F
testDF.los.rf.bt={}
outcomeVar="losBinaryFactor"
for(i in 1:length(data.splitted.los)){
# for(i in 1:5){
  print(i)
  idx=paste0("iter",i)
  
  if(upsampling){
    data=upSampler(data.splitted.los[[idx]]$data2.train.RF,"race3",0.7)
  }else{
    data=data.splitted.los[[idx]]$data2.train.RF
  }
 
  model.rb.rf.ss <- randomForest(formula=rb.eq, 
    data=data, proximity=F,mtry=4) 
  
  model.rn.rf.ss <- randomForest(formula=rn.eq, 
    data=data, proximity=F,mtry=4) 
  
  dfResults=data.splitted.los[[idx]]$data2.valid.RF %>% dplyr::select(race3,losBinaryFactor)
  dfResults$r30d.pred.rb.rf.ss <- factor(predict(model.rb.rf.ss, data.splitted.los[[idx]]$data2.valid.RF),levels=myLevels)
  dfResults$r30d.pred.rn.rf.ss <- factor(predict(model.rn.rf.ss, data.splitted.los[[idx]]$data2.valid.RF),levels=myLevels)
  ######
  # rb all
  cm<-confusionMatrix(
    data=dfResults$r30d.pred.rb.rf.ss,
    reference=dfResults %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                      fairmetrics2DF(F,"All",model1Label,cm))
  # rb black aa
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'Black_AA') %>% pull(r30d.pred.rb.rf.ss), 
    reference = dfResults %>% filter(race3 == 'Black_AA') %>% pull(outcomeVar))
  # print(cm$table)
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"Black_AA",model1Label,cm))
  # rb white ea
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'White_EA') %>% pull(r30d.pred.rb.rf.ss), 
    reference = dfResults %>% filter(race3 == 'White_EA')  %>% pull(outcomeVar))
  # print(cm$table)
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"White_EA",model1Label,cm))
  # rb asian 
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'Asian') %>% pull(r30d.pred.rb.rf.ss), 
    reference = dfResults %>% filter(race3 == 'Asian') %>% pull(outcomeVar))
  # print(cm$table)
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"Asian",model1Label,cm))
  # rb HL
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'HL') %>% pull(r30d.pred.rb.rf.ss), 
    reference = dfResults %>% filter(race3 == 'HL') %>% pull(outcomeVar))
  # print(cm$table)
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"HL",model1Label,cm))
  # print(cm$table)
  # rb MENA
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'MENA') %>% pull(r30d.pred.rb.rf.ss), 
    reference = dfResults %>% filter(race3 == 'MENA') %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"MENA",model1Label,cm))
  
  ######
  # rn all
  cm<-confusionMatrix(
    data=dfResults$r30d.pred.rn.rf.ss,
    reference=dfResults %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(F,"All",model2Label,cm))
  # rn black aa 
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'Black_AA') %>% pull(r30d.pred.rn.rf.ss), 
    reference = dfResults %>% filter(race3 == 'Black_AA') %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"Black_AA",model2Label,cm))
  # rn white ea
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'White_EA') %>% pull(r30d.pred.rn.rf.ss), 
    reference = dfResults %>% filter(race3 == 'White_EA') %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"White_EA",model2Label,cm))
  # rn asian 
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'Asian') %>% pull(r30d.pred.rn.rf.ss), 
    reference = dfResults %>% filter(race3 == 'Asian') %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"Asian",model2Label,cm))
  # rn HL
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'HL') %>% pull(r30d.pred.rn.rf.ss), 
    reference = dfResults %>% filter(race3 == 'HL') %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"HL",model2Label,cm))
  # rn MENA
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'MENA') %>% pull(r30d.pred.rn.rf.ss), 
    reference = dfResults %>% filter(race3 == 'MENA') %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"MENA",model2Label,cm))
}
```


### Random forest  upsampling

```{r}
model1Label="rb.rf.up"
model2Label="rn.rf.up"
upsampling=T
# testDF.los.rf.bt={}
for(i in 1:length(data.splitted.los)){
# for(i in 1:5){
  print(i)
  idx=paste0("iter",i)
  
  if(upsampling){
    data=upSampler(data.splitted.los[[idx]]$data2.train.RF,"race3",0.7)
  }else{
    data=data.splitted.los[[idx]]$data2.train.RF
  }
 
  model.rb.rf.ss <- randomForest(formula=rb.eq, 
    data=data, proximity=F,mtry=4) 
  
  model.rn.rf.ss <- randomForest(formula=rn.eq, 
    data=data, proximity=F,mtry=4) 
  
  dfResults=data.splitted.los[[idx]]$data2.valid.RF %>% dplyr::select(race3,losBinaryFactor)
  dfResults$r30d.pred.rb.rf.ss <- factor(predict(model.rb.rf.ss, data.splitted.los[[idx]]$data2.valid.RF),levels=myLevels)
  dfResults$r30d.pred.rn.rf.ss <- factor(predict(model.rn.rf.ss, data.splitted.los[[idx]]$data2.valid.RF),levels=myLevels)
  ######
  # rb all
  cm<-confusionMatrix(
    data=dfResults$r30d.pred.rb.rf.ss,
    reference=dfResults %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                      fairmetrics2DF(F,"All",model1Label,cm))
  # rb black aa
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'Black_AA') %>% pull(r30d.pred.rb.rf.ss), 
    reference = dfResults %>% filter(race3 == 'Black_AA') %>% pull(outcomeVar))
  # print(cm$table)
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"Black_AA",model1Label,cm))
  # rb white ea
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'White_EA') %>% pull(r30d.pred.rb.rf.ss), 
    reference = dfResults %>% filter(race3 == 'White_EA')  %>% pull(outcomeVar))
  # print(cm$table)
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"White_EA",model1Label,cm))
  # rb asian 
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'Asian') %>% pull(r30d.pred.rb.rf.ss), 
    reference = dfResults %>% filter(race3 == 'Asian') %>% pull(outcomeVar))
  # print(cm$table)
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"Asian",model1Label,cm))
  # rb HL
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'HL') %>% pull(r30d.pred.rb.rf.ss), 
    reference = dfResults %>% filter(race3 == 'HL') %>% pull(outcomeVar))
  # print(cm$table)
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"HL",model1Label,cm))
  # print(cm$table)
  # rb MENA
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'MENA') %>% pull(r30d.pred.rb.rf.ss), 
    reference = dfResults %>% filter(race3 == 'MENA') %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"MENA",model1Label,cm))
  
  ######
  # rn all
  cm<-confusionMatrix(
    data=dfResults$r30d.pred.rn.rf.ss,
    reference=dfResults %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(F,"All",model2Label,cm))
  # rn black aa 
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'Black_AA') %>% pull(r30d.pred.rn.rf.ss), 
    reference = dfResults %>% filter(race3 == 'Black_AA') %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"Black_AA",model2Label,cm))
  # rn white ea
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'White_EA') %>% pull(r30d.pred.rn.rf.ss), 
    reference = dfResults %>% filter(race3 == 'White_EA') %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"White_EA",model2Label,cm))
  # rn asian 
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'Asian') %>% pull(r30d.pred.rn.rf.ss), 
    reference = dfResults %>% filter(race3 == 'Asian') %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"Asian",model2Label,cm))
  # rn HL
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'HL') %>% pull(r30d.pred.rn.rf.ss), 
    reference = dfResults %>% filter(race3 == 'HL') %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"HL",model2Label,cm))
  # rn MENA
  cm<-confusionMatrix(
    data = dfResults %>% filter(race3 == 'MENA') %>% pull(r30d.pred.rn.rf.ss), 
    reference = dfResults %>% filter(race3 == 'MENA') %>% pull(outcomeVar))
  testDF.los.rf.bt<-rbind(testDF.los.rf.bt,
                   fairmetrics2DF(T,"MENA",model2Label,cm))
}
```


```{r}
save(testDF.los.rf.bt,file = "testDF.los.rf.bt.v1.RData",compress = T)
```



-Married and get h4 and then apply 6 months (preferred but not possible)
-Follow to join 1 year (the most suitable)
-3-5 years if filled after getting greencard (i30), she needs her full visa (spouse of a permanent resident). place in line starts after I've got greencard.

Costs for 1 person
Fee for the goverment i485 $1440, 
EAD $260 (optional, it does not hurt)
Advance parole $630 (traveling document), because you are on h1b, having advance parole is a protection just in case if children fires you.
Lawyer $1500 (50:50)

Documents
passport
i94
visa h1b
passport photos
birth certificate, offiiliated translated

US civil surgeon, they give you a seal envelope. ($300ish)

Costs for couple
i24 form $599
visa $445

U6W*NPbEj@.Djih


```{r}
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='White_EA' & algo=='rb.rf') %>% pull(accuracy),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='White_EA' & algo=='rn.rf') %>% pull(accuracy),0.95)

confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='Black_AA' & algo=='rb.rf') %>% pull(accuracy),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='Black_AA' & algo=='rn.rf') %>% pull(accuracy),0.95)

confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='Asian' & algo=='rb.rf') %>% pull(accuracy),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='Asian' & algo=='rn.rf') %>% pull(accuracy),0.95)

confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='HL' & algo=='rb.rf') %>% pull(accuracy),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='HL' & algo=='rn.rf') %>% pull(accuracy),0.95)

confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='MENA' & algo=='rb.rf') %>% pull(accuracy),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='MENA' & algo=='rn.rf') %>% pull(accuracy),0.95)
```

```{r}
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='White_EA' & algo=='rb.rf') %>% pull(mcc),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='White_EA' & algo=='rn.rf') %>% pull(mcc),0.95)

confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='Black_AA' & algo=='rb.rf') %>% pull(mcc),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='Black_AA' & algo=='rn.rf') %>% pull(mcc),0.95)

confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='Asian' & algo=='rb.rf') %>% pull(mcc),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='Asian' & algo=='rn.rf') %>% pull(mcc),0.95)

confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='HL' & algo=='rb.rf') %>% pull(mcc),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='HL' & algo=='rn.rf') %>% pull(mcc),0.95)

confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='MENA' & algo=='rb.rf') %>% pull(mcc),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='MENA' & algo=='rn.rf') %>% pull(mcc),0.95)
```

```{r}
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='White_EA' & algo=='rb.rf') %>% pull(mcc),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='White_EA' & algo=='rn.rf') %>% pull(mcc),0.95)

confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='Black_AA' & algo=='rb.rf') %>% pull(mcc),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='Black_AA' & algo=='rn.rf') %>% pull(mcc),0.95)

confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='Asian' & algo=='rb.rf') %>% pull(mcc),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='Asian' & algo=='rn.rf') %>% pull(mcc),0.95)

confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='HL' & algo=='rb.rf') %>% pull(mcc),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='HL' & algo=='rn.rf') %>% pull(mcc),0.95)

confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='MENA' & algo=='rb.rf') %>% pull(mcc),0.95)
confidence_interval(testDF.los.rf.bt %>% filter(groupoi=='MENA' & algo=='rn.rf') %>% pull(mcc),0.95)
```