---
title: ''
output: html_document
date: "2023-10-10"
---

v1
vis for glm models

v2 
vis for all models including upsampling
includes heatmap and color bar between algorithms

v3 
visualizing only top performers in each race group
color contrast improved
deep code clean up

v4
graphical abstract figures added


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning = F,results = T)
rm(list = ls())
library(tidyverse)
library(patchwork)
library(rstanarm)
library(bayesplot)
library(loo)
library(lme4)
library(gtsummary)
library(fairness)
library(caret)
library(randomForest)
library(xgboost)
library(viridis)

cbPalette <- c("#E64B35", "#4DBBD5")
source("../allofus_los_Jtrans2024/helpers.R")
source("myFairness.R")

myLevels<-c('Yes','No')


get_legend <- function(a.gplot){ 
    tmp <- ggplot_gtable(ggplot_build(a.gplot)) 
    leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box") 
    legend <- tmp$grobs[[leg]] 
    legend
} 
```

```{r}
load(file="testDF.los.glm.bt.v1.RData")
summary(factor(testDF.los.glm.bt$algo))
load(file="testDF.los.rf.bt.v1.RData")
summary(factor(testDF.los.rf.bt$algo))
load(file="testDF.los.xgb.bt.v1.RData")
summary(factor(testDF.los.xgb.bt$algo))
load(file="testDF.los.catb.bt.RData")
summary(factor(testDF.los.catb.bt$algo))

#readmission
load(file = "testDF.read.glm.bt.v3.RData")
summary(factor(testDF.read.glm.bt$algo))
load(file = "testDF.read.rf.bt.v3.RData")
load(file = "testDF.read.xgb.bt.v3.RData")
load(file = "testDF.read.catb.bt.v2.RData")

```





Notes
- If we are looking to sell race-neutral approaches, don't bring additional rb approaches into the conversation.
- Always compare others vs race-based approaches


```{r}
metrics_df_f1<-{}

metrics_df_f1$white.los.auc.base<-testDF.los.glm.bt %>% filter(groupoi=="White_EA" & algo=="rb.glm") %>% pull(accuracy)
metrics_df_f1$white.los.sens.base<-testDF.los.glm.bt %>% filter(groupoi=="White_EA" & algo=="rb.glm") %>% pull(sensitivity)
metrics_df_f1$white.los.spec.base<-testDF.los.glm.bt %>% filter(groupoi=="White_EA" & algo=="rb.glm") %>% pull(specificity)
metrics_df_f1$white.los.pp.base<-testDF.los.glm.bt %>% filter(groupoi=="White_EA" & algo=="rb.glm") %>% pull(prop.parity)
metrics_df_f1$white.los.auc.best<-testDF.los.glm.bt %>% filter(groupoi=="White_EA" & algo=="rb.glm.up") %>% pull(accuracy)
metrics_df_f1$white.los.sens.best<-testDF.los.glm.bt %>% filter(groupoi=="White_EA" & algo=="rb.glm.up") %>% pull(sensitivity)
metrics_df_f1$white.los.spec.best<-testDF.los.glm.bt %>% filter(groupoi=="White_EA" & algo=="rb.glm.up") %>% pull(specificity)
metrics_df_f1$white.los.pp.best<-testDF.los.glm.bt %>% filter(groupoi=="White_EA" & algo=="rb.glm.up") %>% pull(prop.parity)


metrics_df_f1$black.los.auc.base<-testDF.los.glm.bt %>% filter(groupoi=="Black_AA" & algo=="rb.glm") %>% pull(accuracy)
metrics_df_f1$black.los.sens.base<-testDF.los.glm.bt %>% filter(groupoi=="Black_AA" & algo=="rb.glm") %>% pull(sensitivity)
metrics_df_f1$black.los.spec.base<-testDF.los.glm.bt %>% filter(groupoi=="Black_AA" & algo=="rb.glm") %>% pull(specificity)
metrics_df_f1$black.los.pp.base<-testDF.los.glm.bt %>% filter(groupoi=="Black_AA" & algo=="rb.glm") %>% pull(prop.parity)
metrics_df_f1$black.los.auc.best<-testDF.los.glm.bt %>% filter(groupoi=="Black_AA" & algo=="rn.glm.up") %>% pull(accuracy)
metrics_df_f1$black.los.sens.best<-testDF.los.glm.bt %>% filter(groupoi=="Black_AA" & algo=="rn.glm.up") %>% pull(sensitivity)
metrics_df_f1$black.los.spec.best<-testDF.los.glm.bt %>% filter(groupoi=="Black_AA" & algo=="rn.glm.up") %>% pull(specificity)
metrics_df_f1$black.los.pp.best<-testDF.los.glm.bt %>% filter(groupoi=="Black_AA" & algo=="rn.glm.up") %>% pull(prop.parity)

metrics_df_f1$asian.los.auc.base<-testDF.los.glm.bt %>% filter(groupoi=="Asian" & algo=="rb.glm") %>% pull(accuracy)
metrics_df_f1$asian.los.sens.base<-testDF.los.glm.bt %>% filter(groupoi=="Asian" & algo=="rb.glm") %>% pull(sensitivity)
metrics_df_f1$asian.los.spec.base<-testDF.los.glm.bt %>% filter(groupoi=="Asian" & algo=="rb.glm") %>% pull(specificity)
metrics_df_f1$asian.los.pp.base<-testDF.los.glm.bt %>% filter(groupoi=="Asian" & algo=="rb.glm") %>% pull(prop.parity)
metrics_df_f1$asian.los.auc.best<-testDF.los.glm.bt %>% filter(groupoi=="Asian" & algo=="rb.glm.up") %>% pull(accuracy)
metrics_df_f1$asian.los.sens.best<-testDF.los.glm.bt %>% filter(groupoi=="Asian" & algo=="rb.glm.up") %>% pull(sensitivity)
metrics_df_f1$asian.los.spec.best<-testDF.los.glm.bt %>% filter(groupoi=="Asian" & algo=="rb.glm.up") %>% pull(specificity)
metrics_df_f1$asian.los.pp.best<-testDF.los.glm.bt %>% filter(groupoi=="Asian" & algo=="rb.glm.up") %>% pull(prop.parity)

metrics_df_f1$hl.los.auc.base<-testDF.los.glm.bt %>% filter(groupoi=="HL" & algo=="rb.glm") %>% pull(accuracy)
metrics_df_f1$hl.los.sens.base<-testDF.los.glm.bt %>% filter(groupoi=="HL" & algo=="rb.glm") %>% pull(sensitivity)
metrics_df_f1$hl.los.spec.base<-testDF.los.glm.bt %>% filter(groupoi=="HL" & algo=="rb.glm") %>% pull(specificity)
metrics_df_f1$hl.los.pp.base<-testDF.los.glm.bt %>% filter(groupoi=="HL" & algo=="rb.glm") %>% pull(prop.parity)
metrics_df_f1$hl.los.auc.best<-testDF.los.glm.bt %>% filter(groupoi=="HL" & algo=="rn.glm.up") %>% pull(accuracy)
metrics_df_f1$hl.los.sens.best<-testDF.los.glm.bt %>% filter(groupoi=="HL" & algo=="rn.glm.up") %>% pull(sensitivity)
metrics_df_f1$hl.los.spec.best<-testDF.los.glm.bt %>% filter(groupoi=="HL" & algo=="rn.glm.up") %>% pull(specificity)
metrics_df_f1$hl.los.pp.best<-testDF.los.glm.bt %>% filter(groupoi=="HL" & algo=="rn.glm.up") %>% pull(prop.parity)

metrics_df_f1$mena.los.auc.base<-testDF.los.glm.bt %>% filter(groupoi=="MENA" & algo=="rb.glm") %>% pull(accuracy)
metrics_df_f1$mena.los.sens.base<-testDF.los.glm.bt %>% filter(groupoi=="MENA" & algo=="rb.glm") %>% pull(sensitivity)
metrics_df_f1$mena.los.spec.base<-testDF.los.glm.bt %>% filter(groupoi=="MENA" & algo=="rb.glm") %>% pull(specificity)
metrics_df_f1$mena.los.pp.base<-testDF.los.glm.bt %>% filter(groupoi=="MENA" & algo=="rb.glm") %>% pull(prop.parity)
metrics_df_f1$mena.los.auc.best<-testDF.los.glm.bt %>% filter(groupoi=="MENA" & algo=="rb.glm.up") %>% pull(accuracy)
metrics_df_f1$mena.los.sens.best<-testDF.los.glm.bt %>% filter(groupoi=="MENA" & algo=="rb.glm.up") %>% pull(sensitivity)
metrics_df_f1$mena.los.spec.best<-testDF.los.glm.bt %>% filter(groupoi=="MENA" & algo=="rb.glm.up") %>% pull(specificity)
metrics_df_f1$mena.los.pp.best<-testDF.los.glm.bt %>% filter(groupoi=="MENA" & algo=="rb.glm.up") %>% pull(prop.parity)

## Readmission
metrics_df_f1$white.read.auc.base<-testDF.read.glm.bt %>% filter(groupoi=="White_EA" & algo=="rb.glm") %>% pull(accuracy)
metrics_df_f1$white.read.sens.base<-testDF.read.glm.bt %>% filter(groupoi=="White_EA" & algo=="rb.glm") %>% pull(sensitivity)
metrics_df_f1$white.read.spec.base<-testDF.read.glm.bt %>% filter(groupoi=="White_EA" & algo=="rb.glm") %>% pull(specificity)
metrics_df_f1$white.read.pp.base<-testDF.read.glm.bt %>% filter(groupoi=="White_EA" & algo=="rb.glm") %>% pull(prop.parity)
metrics_df_f1$white.read.auc.best<-testDF.read.xgb.bt %>% filter(groupoi=="White_EA" & algo=="rn.xgb") %>% pull(accuracy)
metrics_df_f1$white.read.sens.best<-testDF.read.xgb.bt %>% filter(groupoi=="White_EA" & algo=="rn.xgb") %>% pull(sensitivity)
metrics_df_f1$white.read.spec.best<-testDF.read.xgb.bt %>% filter(groupoi=="White_EA" & algo=="rn.xgb") %>% pull(specificity)
metrics_df_f1$white.read.pp.best<-testDF.read.xgb.bt %>% filter(groupoi=="White_EA" & algo=="rn.xgb") %>% pull(prop.parity)

metrics_df_f1$black.read.auc.base<-testDF.read.glm.bt %>% filter(groupoi=="Black_AA" & algo=="rb.glm") %>% pull(accuracy)
metrics_df_f1$black.read.sens.base<-testDF.read.glm.bt %>% filter(groupoi=="Black_AA" & algo=="rb.glm") %>% pull(sensitivity)
metrics_df_f1$black.read.spec.base<-testDF.read.glm.bt %>% filter(groupoi=="Black_AA" & algo=="rb.glm") %>% pull(specificity)
metrics_df_f1$black.read.pp.base<-testDF.read.glm.bt %>% filter(groupoi=="Black_AA" & algo=="rb.glm") %>% pull(prop.parity)
metrics_df_f1$black.read.auc.best<-testDF.read.xgb.bt %>% filter(groupoi=="Black_AA" & algo=="rb.xgb") %>% pull(accuracy)
metrics_df_f1$black.read.sens.best<-testDF.read.xgb.bt %>% filter(groupoi=="Black_AA" & algo=="rb.xgb") %>% pull(sensitivity)
metrics_df_f1$black.read.spec.best<-testDF.read.xgb.bt %>% filter(groupoi=="Black_AA" & algo=="rb.xgb") %>% pull(specificity)
metrics_df_f1$black.read.pp.best<-testDF.read.xgb.bt %>% filter(groupoi=="Black_AA" & algo=="rb.xgb") %>% pull(prop.parity)

metrics_df_f1$asian.read.auc.base<-testDF.read.glm.bt %>% filter(groupoi=="Asian" & algo=="rb.glm") %>% pull(accuracy)
metrics_df_f1$asian.read.sens.base<-testDF.read.glm.bt %>% filter(groupoi=="Asian" & algo=="rb.glm") %>% pull(sensitivity)
metrics_df_f1$asian.read.spec.base<-testDF.read.glm.bt %>% filter(groupoi=="Asian" & algo=="rb.glm") %>% pull(specificity)
metrics_df_f1$asian.read.pp.base<-testDF.read.glm.bt %>% filter(groupoi=="Asian" & algo=="rb.glm") %>% pull(prop.parity)
metrics_df_f1$asian.read.auc.best<-testDF.read.xgb.bt %>% filter(groupoi=="Asian" & algo=="rb.xgb") %>% pull(accuracy)
metrics_df_f1$asian.read.sens.best<-testDF.read.xgb.bt %>% filter(groupoi=="Asian" & algo=="rb.xgb") %>% pull(sensitivity)
metrics_df_f1$asian.read.spec.best<-testDF.read.xgb.bt %>% filter(groupoi=="Asian" & algo=="rb.xgb") %>% pull(specificity)
metrics_df_f1$asian.read.pp.best<-testDF.read.xgb.bt %>% filter(groupoi=="Asian" & algo=="rb.xgb") %>% pull(prop.parity)

metrics_df_f1$hl.read.auc.base<-testDF.read.glm.bt %>% filter(groupoi=="HL" & algo=="rb.glm") %>% pull(accuracy)
metrics_df_f1$hl.read.sens.base<-testDF.read.glm.bt %>% filter(groupoi=="HL" & algo=="rb.glm") %>% pull(sensitivity)
metrics_df_f1$hl.read.spec.base<-testDF.read.glm.bt %>% filter(groupoi=="HL" & algo=="rb.glm") %>% pull(specificity)
metrics_df_f1$hl.read.pp.base<-testDF.read.glm.bt %>% filter(groupoi=="HL" & algo=="rb.glm") %>% pull(prop.parity)
metrics_df_f1$hl.read.auc.best<-testDF.read.xgb.bt %>% filter(groupoi=="HL" & algo=="rn.xgb") %>% pull(accuracy)
metrics_df_f1$hl.read.sens.best<-testDF.read.xgb.bt %>% filter(groupoi=="HL" & algo=="rn.xgb") %>% pull(sensitivity)
metrics_df_f1$hl.read.spec.best<-testDF.read.xgb.bt %>% filter(groupoi=="HL" & algo=="rn.xgb") %>% pull(specificity)
metrics_df_f1$hl.read.pp.best<-testDF.read.xgb.bt %>% filter(groupoi=="HL" & algo=="rn.xgb") %>% pull(prop.parity)

metrics_df_f1$mena.read.auc.base<-testDF.read.glm.bt %>% filter(groupoi=="MENA" & algo=="rb.glm") %>% pull(accuracy)
metrics_df_f1$mena.read.sens.base<-testDF.read.glm.bt %>% filter(groupoi=="MENA" & algo=="rb.glm") %>% pull(sensitivity)
metrics_df_f1$mena.read.spec.base<-testDF.read.glm.bt %>% filter(groupoi=="MENA" & algo=="rb.glm") %>% pull(specificity)
metrics_df_f1$mena.read.pp.base<-testDF.read.glm.bt %>% filter(groupoi=="MENA" & algo=="rb.glm") %>% pull(prop.parity)
metrics_df_f1$mena.read.auc.best<-testDF.read.xgb.bt %>% filter(groupoi=="MENA" & algo=="rn.xgb") %>% pull(accuracy)
metrics_df_f1$mena.read.sens.best<-testDF.read.xgb.bt %>% filter(groupoi=="MENA" & algo=="rn.xgb") %>% pull(sensitivity)
metrics_df_f1$mena.read.spec.best<-testDF.read.xgb.bt %>% filter(groupoi=="MENA" & algo=="rn.xgb") %>% pull(specificity)
metrics_df_f1$mena.read.pp.best<-testDF.read.xgb.bt %>% filter(groupoi=="MENA" & algo=="rn.xgb") %>% pull(prop.parity)

metrics_df_f1<-as.data.frame(metrics_df_f1)
summary(metrics_df_f1)

metrics_df_f1_long<-metrics_df_f1 %>%
  # dplyr::select(ratio_wb,ratio_wa,ratio_wh,ratio_wm) %>%
  gather(.,metric,value,white.los.auc.base:mena.read.pp.best,factor_key = T)%>% 
  mutate(group="Baseline") %>% 
  separate_wider_delim(metric, delim = ".", names = c("groupoi","outcome", "metric","model"))

# summary(factor(metrics_df_f1_long$metric))

(df_f1_summary<-metrics_df_f1_long %>% 
  group_by(outcome,groupoi,metric,model) %>% 
  summarise(mean=mean(value),
            mysd=sd(value),
            mylength=length(value),
            myse=mysd/sqrt(mylength),
            ci=myse*qt((1-0.05)/2+0.5,mylength-1)
            ) 
)

df_f1_summary$groupoi<-factor(df_f1_summary$groupoi,levels = (c('white','black','asian','hl','mena')))
df_f1_summary$model<-factor(df_f1_summary$model,levels = (c('base','best')))
df_f1_summary$outcome<-factor(df_f1_summary$outcome,levels = (c('los','read')))
df_f1_summary$metric<-factor(df_f1_summary$metric,levels = (c('auc','sens','spec','pp')))

```

```{r}
facet_names <- c(
  `los` = "LoS",
  `read` = "Readmission",
  `white` = "White",
                 `black` = "Black",
                 `asian` = "Asian",
                 `hl` = "HL",
                 `mena` = "MENA"
                )

xlabels=c("AUC","Sensitivity","Specificity")
flabels=c("Baseline","Best model")
(f1a<-df_f1_summary%>%
  filter(outcome=="los" & metric!="pp") %>% 
  ggplot(aes(x = metric, y = mean, fill = model,ymin=mean-ci,ymax=mean+ci)) +
  # geom_col(color = "black", width = 0.9,position = "dodge2")+
  geom_bar(stat = "identity",position = position_dodge2(),alpha=0.7)+
  geom_errorbar(position = position_dodge(width=0.9),width=0.4,colour="black",alpha=0.9,size=0.5)+
  coord_cartesian(ylim=c(0, 1))+
  scale_x_discrete(guide = guide_axis(angle=0,n.dodge = 2)
                   ,labels=xlabels
                   # ,limits=rev(levels(ratio))
                   ) +
  # scale_fill_brewer(palette = "Greens"
  #                   ,labels=flabels
  #                   )+
    scale_fill_manual(values=c("#69B41E","#0D5B11")
                      ,labels=flabels
                      )+
  labs(subtitle="A) Asthma LoS",fill="LoS Model",y="Score",x="")+
  facet_wrap(vars(groupoi),labeller=as_labeller(facet_names),ncol = 5)+
  theme_classic()
)


(f1b<-df_f1_summary%>%
  filter(outcome=="read" & metric!="pp") %>% 
  ggplot(aes(x = metric, y = mean, fill = model,ymin=mean-ci,ymax=mean+ci)) +
  # geom_col(color = "black", width = 0.9,position = "dodge2")+
  geom_bar(stat = "identity",position = position_dodge2(),alpha=0.7)+
  geom_errorbar(position = position_dodge(width=0.9),width=0.4,colour="black",alpha=0.9,size=0.5)+
  coord_cartesian(ylim=c(0, 1))+
  scale_x_discrete(guide = guide_axis(angle=0,n.dodge = 2)
                   ,labels=xlabels
                   # ,limits=rev(levels(ratio))
                   ) +
    
  # scale_fill_brewer(palette = "Blues"
  #                   ,labels=flabels
  #                   )+
    scale_fill_manual(values=c("#4169E1","#031279")
                      ,labels=flabels)+
  labs(subtitle="B) Asthma Readmission",fill="Readmission\nModel",y="Score",x="")+
  facet_wrap(vars(groupoi),labeller=as_labeller(facet_names),ncol = 5)+
  theme_classic()
)

(f1.full<-f1a/f1b)

# ggsave(filename = "images/f1.v3b.png", f1.full,# en cm mejor para mantener el tamanho
# width=3400,height = 2200, dpi = 300, units = "px", device='png')
```


## GA

```{r}
xlabels=c("AUC","Sensitivity","Specificity","Prop. Parity")
flabels=c("Baseline")

(ga1<-df_f1_summary%>%
  filter(outcome=="read" & model=="base") %>% 
  ggplot(aes(x = metric, y = mean, fill = model,ymin=mean-ci,ymax=mean+ci)) +
  # geom_col(color = "black", width = 0.9,position = "dodge2")+
  geom_bar(stat = "identity",position = position_dodge2(),alpha=0.7)+
  geom_errorbar(position = position_dodge(width=0.9),width=0.4,colour="black",alpha=0.9,size=0.5)+
  coord_cartesian(ylim=c(0, 1))+
  scale_x_discrete(guide = guide_axis(angle=45
                                      # ,n.dodge = 0
                                      )
                   ,labels=xlabels
                   ) +

    scale_fill_manual(values=c("#4169E1","#031279")
                      ,labels=flabels)+
  labs(subtitle="Baseline Model",fill="Readmission\nModel",y="Score",x="")+
  facet_wrap(vars(groupoi),labeller=as_labeller(facet_names),ncol = 5)+
  theme_classic()+
    theme(legend.position = "None")
)
flabels=c("Best model")
(ga2<-df_f1_summary%>%
  filter(outcome=="read" & model=="best") %>% 
  ggplot(aes(x = metric, y = mean, fill = model,ymin=mean-ci,ymax=mean+ci)) +
  geom_bar(stat = "identity",position = position_dodge2(),alpha=0.7)+
  geom_errorbar(position = position_dodge(width=0.9),width=0.4,colour="black",alpha=0.9,size=0.5)+
  coord_cartesian(ylim=c(0, 1))+
  scale_x_discrete(guide = guide_axis(angle=45
                                      # ,n.dodge = 2
                                      )
                   ,labels=xlabels
                   ) +

    scale_fill_manual(values=c("#031279")
                      ,labels=flabels)+
  labs(subtitle="Debiased Model",fill="Readmission\nModel",y="Score",x="")+
  facet_wrap(vars(groupoi),labeller=as_labeller(facet_names),ncol = 5)+
  theme_classic()+
    theme(legend.position = "None")
)

(ga.full<-ga1+ga2)

ggsave(filename = "images/ga.v1.png", ga.full,# en cm mejor para mantener el tamanho
width=3400,height = 2200, dpi = 300, units = "px", device='png')
```


## p-values for los

between non-white and reference using top algorithms

```{r}
t.test(metrics_df_f1$black.los.auc.best, metrics_df_f1$white.los.auc.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$asian.los.auc.best, metrics_df_f1$white.los.auc.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$hl.los.auc.best, metrics_df_f1$white.los.auc.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$mena.los.auc.best, metrics_df_f1$white.los.auc.best, alternative = "two.sided", var.equal = FALSE)

t.test(metrics_df_f1$black.los.sens.best, metrics_df_f1$white.los.sens.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$asian.los.sens.best, metrics_df_f1$white.los.sens.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$hl.los.sens.best, metrics_df_f1$white.los.sens.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$mena.los.sens.best, metrics_df_f1$white.los.sens.best, alternative = "two.sided", var.equal = FALSE)

t.test(metrics_df_f1$black.los.spec.best, metrics_df_f1$white.los.spec.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$asian.los.spec.best, metrics_df_f1$white.los.spec.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$hl.los.spec.best, metrics_df_f1$white.los.spec.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$mena.los.spec.best, metrics_df_f1$white.los.spec.best, alternative = "two.sided", var.equal = FALSE)

t.test(metrics_df_f1$black.los.pp.best, metrics_df_f1$white.los.pp.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$asian.los.pp.best, metrics_df_f1$white.los.pp.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$hl.los.pp.best, metrics_df_f1$white.los.pp.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$mena.los.pp.best, metrics_df_f1$white.los.pp.best, alternative = "two.sided", var.equal = FALSE)
```
## p-values for read

between non-white and reference using top algorithms

```{r}
t.test(metrics_df_f1$black.read.auc.best, metrics_df_f1$white.read.auc.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$asian.read.auc.best, metrics_df_f1$white.read.auc.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$hl.read.auc.best, metrics_df_f1$white.read.auc.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$mena.read.auc.best, metrics_df_f1$white.read.auc.best, alternative = "two.sided", var.equal = FALSE)

t.test(metrics_df_f1$black.read.sens.best, metrics_df_f1$white.read.sens.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$asian.read.sens.best, metrics_df_f1$white.read.sens.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$hl.read.sens.best, metrics_df_f1$white.read.sens.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$mena.read.sens.best, metrics_df_f1$white.read.sens.best, alternative = "two.sided", var.equal = FALSE)

t.test(metrics_df_f1$black.read.spec.best, metrics_df_f1$white.read.spec.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$asian.read.spec.best, metrics_df_f1$white.read.spec.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$hl.read.spec.best, metrics_df_f1$white.read.spec.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$mena.read.spec.best, metrics_df_f1$white.read.spec.best, alternative = "two.sided", var.equal = FALSE)

t.test(metrics_df_f1$black.read.pp.best, metrics_df_f1$white.read.pp.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$asian.read.pp.best, metrics_df_f1$white.read.pp.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$hl.read.pp.best, metrics_df_f1$white.read.pp.best, alternative = "two.sided", var.equal = FALSE)
t.test(metrics_df_f1$mena.read.pp.best, metrics_df_f1$white.read.pp.best, alternative = "two.sided", var.equal = FALSE)
```


# LoS

## ratio unfairness baseline

```{r}
lratio=c('Black vs White','Asian vs White','Hl vs White','MENA vs White')
di_df_los_baseline<-{}

di_df_los_baseline$ratio_wb<-metrics_df_f1$black.los.pp.base/metrics_df_f1$white.los.pp.base
di_df_los_baseline$ratio_wa<-metrics_df_f1$asian.los.pp.base/metrics_df_f1$white.los.pp.base
di_df_los_baseline$ratio_wh<-metrics_df_f1$hl.los.pp.base/metrics_df_f1$white.los.pp.base
di_df_los_baseline$ratio_wm<-metrics_df_f1$mena.los.pp.base/metrics_df_f1$white.los.pp.base

di_df_los_baseline<-as.data.frame(di_df_los_baseline)
summary(di_df_los_baseline)

quantile(di_df_los_baseline$ratio_wb,probs = c(0.025,0.5,0.975))
quantile(di_df_los_baseline$ratio_wa,probs = c(0.025,0.5,0.975))
quantile(di_df_los_baseline$ratio_wh,probs = c(0.025,0.5,0.975))
quantile(di_df_los_baseline$ratio_wm,probs = c(0.025,0.5,0.975))

di_df_los_baseline2<-di_df_los_baseline %>%
  dplyr::select(ratio_wb,ratio_wa,ratio_wh,ratio_wm) %>%
  gather(.,ratio,value,ratio_wb:ratio_wm,factor_key = T)%>% 
  mutate(group="Baseline")

pp_df_los_baseline<-metrics_df_f1 %>%
  dplyr::select(white.los.pp.base,
                black.los.pp.base,
                asian.los.pp.base,
                hl.los.pp.base,
                mena.los.pp.base
                ) %>%
  gather(.,pp,value,white.los.pp.base:mena.los.pp.base,factor_key = T)%>% 
  mutate(group="Baseline")

ggplot(di_df_los_baseline2,aes(x=ratio,y=value)) + 
  geom_boxplot(aes(fill=ratio),outliers = F)+
  # coord_cartesian(ylim=c(-0.1,0.1))+
  geom_hline(yintercept = 1,lty="dashed",color="blue")+
  stat_summary(fun.y = mean, geom="point",colour="darkred", size=2) +
  scale_x_discrete(guide = guide_axis(angle=75)
                   ,labels=lratio
                   ) +
  scale_fill_viridis_d(option = 8,labels=lratio)+
  labs(fill="Method",y="",x="",subtitle="Readmission's DI Ratio (Baseline)")+
  theme_classic()+
  theme(legend.position = "none")
```

## ratio unfairness after mitigation

```{r}
lratio=c('Black vs White','Asian vs White','HL vs White','MENA vs White')
di_df_los_post<-{}

di_df_los_post$ratio_wb<-metrics_df_f1$black.los.pp.best/metrics_df_f1$white.los.pp.best
di_df_los_post$ratio_wa<-metrics_df_f1$asian.los.pp.best/metrics_df_f1$white.los.pp.best
di_df_los_post$ratio_wh<-metrics_df_f1$hl.los.pp.best/metrics_df_f1$white.los.pp.best
di_df_los_post$ratio_wm<-metrics_df_f1$mena.los.pp.best/metrics_df_f1$white.los.pp.best

di_df_los_post<-as.data.frame(di_df_los_post)
summary(di_df_los_post)

quantile(di_df_los_post$ratio_wb,probs = c(0.025,0.5,0.975))
quantile(di_df_los_post$ratio_wa,probs = c(0.025,0.5,0.975))
quantile(di_df_los_post$ratio_wh,probs = c(0.025,0.5,0.975))
quantile(di_df_los_post$ratio_wm,probs = c(0.025,0.5,0.975))

di_df_los_post2<-di_df_los_post %>%
  dplyr::select(ratio_wb,ratio_wa,ratio_wh,ratio_wm) %>%
  gather(.,ratio,value,ratio_wb:ratio_wm,factor_key = T)%>% 
  mutate(group="Best model")

pp_df_los_post<-metrics_df_f1 %>%
  dplyr::select(white.los.pp.best,
                black.los.pp.best,
                asian.los.pp.best,
                hl.los.pp.best,
                mena.los.pp.best
                ) %>%
  gather(.,pp,value,white.los.pp.best:mena.los.pp.best,factor_key = T)%>% 
  mutate(group="Best model")

ggplot(di_df_los_post2,aes(x=ratio,y=value)) + 
  geom_boxplot(aes(fill=ratio),outliers = F)+
  # coord_cartesian(ylim=c(-0.1,0.1))+
  geom_hline(yintercept = 1,lty="dashed",color="blue")+
  stat_summary(fun.y = mean, geom="point",colour="darkred", size=2) +
  scale_x_discrete(guide = guide_axis(angle=75)
                   ,labels=lratio
                   ) +
  scale_fill_viridis_d(option = 8,labels=lratio)+
  labs(fill="Method",y="",x="",subtitle="Readmission's DI Ratio (Baseline)")+
  theme_classic()+
  theme(legend.position = "none")
```

```{r}
pp_df_los_both=rbind(pp_df_los_baseline,pp_df_los_post) 
pp_summary<-pp_df_los_both %>% 
  group_by(pp,group) %>% 
  summarise(mean=mean(value),
            mysd=sd(value),
            mylength=length(value),
            myse=mysd/sqrt(mylength),
            ci=myse*qt((1-0.05)/2+0.5,mylength-1)
            ) %>% 
  separate_wider_delim(pp, delim = ".", names = c("groupoi","outcome", "metric","model")) %>% 
  mutate(groupoi=factor(groupoi,levels=c("white","black","asian","hl","mena"))
    ,group=factor(group,levels=c("Baseline","Best model")))
pp_summary


lpp=c('White','Black','Asian','HL','MENA')
(a<-ggplot(pp_summary,aes(x=groupoi,y=mean,fill=group,ymin=mean-ci,ymax=mean+ci))+
  geom_bar(stat = "identity",position = position_dodge(),alpha=0.7)+
  geom_errorbar(position = position_dodge(width=0.9),width=0.4,colour="orange",alpha=0.9,size=1.0)+
  scale_x_discrete(guide = guide_axis(angle=0,n.dodge = 2)
                   ,labels=(lpp)
                   # ,limits=rev(levels(pp))
                   ) +
  # scale_fill_brewer(palette = "Greens")+
    scale_fill_manual(values=c("#69B41E","#0D5B11"))+
  labs(fill="LoS Model",y="Prop. Parity (%)",x=""
       ,subtitle="A) Asthma LoS"
       )+
  theme_classic()+
  theme(legend.position = "right"
        # ,panel.background = element_rect(fill = "gray60",
        #                         colour = "gray60",
        #                         size = 0.5, linetype = "solid")
        )
)

legend.los<-get_legend(a)
a<-a+theme(legend.position = "none")

```

```{r}
di_df_los_both=rbind(di_df_los_baseline2,di_df_los_post2) 
(b<-ggplot(di_df_los_both,
           aes(x=ratio,
               y=value,
               fill=factor(group,levels=c('Baseline','Best model')))) + 
  geom_boxplot(outliers = F)+
  geom_hline(yintercept = 1,lty="dashed",color="blue")+
  scale_x_discrete(guide = guide_axis(angle=0,n.dodge = 2)
                   ,labels=(lratio)
                   # ,limits=rev(levels(ratio))
                   ) +
  # scale_fill_brewer(palette = "Greens")+
        scale_fill_manual(values=c("#69B41E","#0D5B11"))+

  labs(fill="Simulations",y="DI (Ratio)",x="")+
  theme_classic()+
  theme(legend.position = "none"
        # ,
        # panel.background = element_rect(fill = "gray60",
        #                         colour = "gray60",
        #                         size = 0.5, linetype = "solid")
        )
)
```

# Readmission

## ratio unfairness baseline

```{r}
lratio=c('Black vs White','Asian vs White','HL vs White','MENA vs White')
di_df_read<-{}

di_df_read$ratio_wb<-metrics_df_f1$black.read.pp.base/metrics_df_f1$white.read.pp.base
di_df_read$ratio_wa<-metrics_df_f1$asian.read.pp.base/metrics_df_f1$white.read.pp.base
di_df_read$ratio_wh<-metrics_df_f1$hl.read.pp.base/metrics_df_f1$white.read.pp.base
di_df_read$ratio_wm<-metrics_df_f1$mena.read.pp.base/metrics_df_f1$white.read.pp.base
di_df_read<-as.data.frame(di_df_read)
summary(di_df_read)

quantile(di_df_read$ratio_wb,probs = c(0.025,0.5,0.975))
quantile(di_df_read$ratio_wa,probs = c(0.025,0.5,0.975))
quantile(di_df_read$ratio_wh,probs = c(0.025,0.5,0.975))
quantile(di_df_read$ratio_wm,probs = c(0.025,0.5,0.975))

di_df_read2<-di_df_read %>%
  dplyr::select(ratio_wb,ratio_wa,ratio_wh,ratio_wm) %>%
  gather(.,ratio,value,ratio_wb:ratio_wm,factor_key = T)%>% 
  mutate(group="Baseline")

pp_df_read_baseline<-metrics_df_f1 %>%
  dplyr::select(white.read.pp.base,
              black.read.pp.base,
              asian.read.pp.base,
              hl.read.pp.base,
              mena.read.pp.base
              ) %>%
  gather(.,pp,value,white.read.pp.base:mena.read.pp.base,factor_key = T)%>% 
  mutate(group="Baseline")

ggplot(di_df_read2,aes(x=ratio,y=value)) + 
  geom_boxplot(aes(fill=ratio),outliers = F)+
  # coord_cartesian(ylim=c(-0.1,0.1))+
  geom_hline(yintercept = 1,lty="dashed",color="blue")+
  stat_summary(fun.y = mean, geom="point",colour="darkred", size=2) +
  scale_x_discrete(guide = guide_axis(angle=75)
                   ,labels=lratio
                   ) +
  scale_fill_viridis_d(option = 8,labels=lratio)+
  labs(fill="Method",y="",x="",subtitle="Readmission's DI Ratio (Baseline)")+
  theme_classic()+
  theme(legend.position = "none")
```


## ratio unfairness after mitigation

```{r}
lratio=c('Black vs White','Asian vs White','HL vs White','MENA vs White')
di_df_read_post<-{}
di_df_read_post$ratio_wb<-metrics_df_f1$black.read.pp.best/metrics_df_f1$white.read.pp.best
di_df_read_post$ratio_wa<-metrics_df_f1$asian.read.pp.best/metrics_df_f1$white.read.pp.best
di_df_read_post$ratio_wh<-metrics_df_f1$hl.read.pp.best/metrics_df_f1$white.read.pp.best
di_df_read_post$ratio_wm<-metrics_df_f1$mena.read.pp.best/metrics_df_f1$white.read.pp.best

di_df_read_post<-as.data.frame(di_df_read_post)
summary(di_df_read_post)

quantile(di_df_read_post$ratio_wb,probs = c(0.025,0.5,0.975))
quantile(di_df_read_post$ratio_wa,probs = c(0.025,0.5,0.975))
quantile(di_df_read_post$ratio_wh,probs = c(0.025,0.5,0.975))
quantile(di_df_read_post$ratio_wm,probs = c(0.025,0.5,0.975))

di_df_read_post2<-di_df_read_post %>%
  dplyr::select(ratio_wb,ratio_wa,ratio_wh,ratio_wm) %>%
  gather(.,ratio,value,ratio_wb:ratio_wm,factor_key = T) %>% 
  mutate(group="Best model")

pp_df_read_post<-metrics_df_f1 %>%
  dplyr::select(white.read.pp.best,
              black.read.pp.best,
              asian.read.pp.best,
              hl.read.pp.best,
              mena.read.pp.best
              ) %>%
  gather(.,pp,value,white.read.pp.best:mena.read.pp.best,factor_key = T)%>% 
  mutate(group="Best model")

ggplot(di_df_read_post2,aes(x=ratio,y=value)) + 
  geom_boxplot(aes(fill=ratio),outliers = F)+
  # coord_cartesian(ylim=c(-0.2,0.1))+
  # coord_flip()+
  geom_hline(yintercept = 1,lty="dashed",color="blue")+
  stat_summary(fun.y = mean, geom="point",colour="darkred", size=2) +
  scale_x_discrete(guide = guide_axis(angle=0)
                   ,labels=lratio
                   ) +
  scale_fill_viridis_d(option = 8,labels=lratio)+
  labs(fill="Method",y="",x="",subtitle="Readmission's DI Ratio (after unfairness mitigation)")+
  theme_classic()+
  theme(legend.position = "none")
```

```{r}
pp_df_both=rbind(pp_df_read_baseline,pp_df_read_post) 
pp_summary<-pp_df_both %>% 
  group_by(pp,group) %>% 
  summarise(mean=mean(value),
            mysd=sd(value),
            mylength=length(value),
            myse=mysd/sqrt(mylength),
            ci=myse*qt((1-0.05)/2+0.5,mylength-1)
            ) %>% 
  separate_wider_delim(pp, delim = ".", names = c("groupoi","outcome", "metric","model")) %>% 
  mutate(groupoi=factor(groupoi,levels=c("white","black","asian","hl","mena"))
    ,group=factor(group,levels=c("Baseline","Best model")))

pp_summary

lpp=c('White','Black','Asian','HL','MENA')
(c<-ggplot(pp_summary,aes(x=groupoi,y=mean,fill=group,ymin=mean-ci,ymax=mean+ci))+
  geom_bar(stat = "identity",position = position_dodge(),alpha=0.7)+
  geom_errorbar(position = position_dodge(width=0.9),width=0.4,colour="orange",alpha=0.9,size=1.0)+
  scale_x_discrete(guide = guide_axis(angle=0,n.dodge = 2)
                   ,labels=(lpp)
                   # ,limits=rev(levels(pp))
                   ) +
  # scale_fill_brewer()+
        scale_fill_manual(values=c("#4169E1","#031279"))+
  labs(fill="Readmission\nModel",y="Prop. Parity (%)",x=""
       , subtitle="B) Asthma Readmission"
       # ,subtitle="Prop. parity across races (before vs after mitigation)"
       )+
  theme_classic()+
  theme(legend.position = "right"
        # ,
        # panel.background = element_rect(fill = "gray60",
        #                         colour = "gray60",
        #                         size = 0.5, linetype = "solid")
        )
)

legendread<-get_legend(c)
c<-c+theme(legend.position = "none")
```

Disparate impact ratio

```{r}
di_df_read_both=rbind(di_df_read2,di_df_read_post2) 
(d<-ggplot(di_df_read_both,aes(
  # x=forcats::fct_rev(ratio),
                      x=ratio
                      ,y=value,fill=factor(group,levels=c('Baseline','Best model')))) + 
  geom_boxplot(outliers = F)+
  geom_hline(yintercept = 1,lty="dashed",color="blue")+
  # stat_summary(fun.y = mean, geom="point",colour="darkred", size=2) +
  scale_x_discrete(guide = guide_axis(angle=0,n.dodge = 2)
                   ,labels=(lratio)
                   # ,limits=rev(levels(ratio))
                   ) +
  # scale_fill_brewer(palette = "Blues")+
            scale_fill_manual(values=c("#4169E1","#031279"))+

  labs(fill="Simulations",y="DI (Ratio)",x=""
       # ,subtitle="DI Ratio (before and after mitigation)"
       )+
  theme_classic()+
  theme(legend.position = "none"
        # ,panel.background = element_rect(fill = "gray60",
        #                         colour = "gray60",
        #                         size = 0.5, linetype = "solid")
        )
)
```





```{r}
(f2.full<-(a+b+legend.los)/(c+d+legendread))

# ggsave(filename = "images/f2.v3b.png", f2.full,# en cm mejor para mantener el tamanho
#   width=2800,height = 2200, dpi = 300, units = "px", device='png')
```




```{r}
t.test(di_df_los_both %>% filter(ratio=="ratio_wb" & group=="Baseline") %>% pull(value),
       di_df_los_both %>% filter(ratio=="ratio_wb" & group=="Best model") %>% pull(value),
       alternative = "two.sided", var.equal = FALSE)

t.test(di_df_los_both %>% filter(ratio=="ratio_wa" & group=="Baseline") %>% pull(value),
       di_df_los_both %>% filter(ratio=="ratio_wa" & group=="Best model") %>% pull(value),
       alternative = "two.sided", var.equal = FALSE)

t.test(di_df_los_both %>% filter(ratio=="ratio_wh" & group=="Baseline") %>% pull(value),
       di_df_los_both %>% filter(ratio=="ratio_wh" & group=="Best model") %>% pull(value),
       alternative = "two.sided", var.equal = FALSE)

t.test(di_df_los_both %>% filter(ratio=="ratio_wm" & group=="Baseline") %>% pull(value),
       di_df_los_both %>% filter(ratio=="ratio_wm" & group=="Best model") %>% pull(value),
       alternative = "two.sided", var.equal = FALSE)
```

```{r}
t.test(di_df_read_both %>% filter(ratio=="ratio_wb" & group=="Baseline") %>% pull(value),
       di_df_read_both %>% filter(ratio=="ratio_wb" & group=="Best model") %>% pull(value),
       alternative = "two.sided", var.equal = FALSE)

t.test(di_df_read_both %>% filter(ratio=="ratio_wa" & group=="Baseline") %>% pull(value),
       di_df_read_both %>% filter(ratio=="ratio_wa" & group=="Best model") %>% pull(value),
       alternative = "two.sided", var.equal = FALSE)

t.test(di_df_read_both %>% filter(ratio=="ratio_wh" & group=="Baseline") %>% pull(value),
       di_df_read_both %>% filter(ratio=="ratio_wh" & group=="Best model") %>% pull(value),
       alternative = "two.sided", var.equal = FALSE)

t.test(di_df_read_both %>% filter(ratio=="ratio_wm" & group=="Baseline") %>% pull(value),
       di_df_read_both %>% filter(ratio=="ratio_wm" & group=="Best model") %>% pull(value),
       alternative = "two.sided", var.equal = FALSE)
```

